Theory about Normal Basis on $F(2^k)$

### 2016-12-11 dissertation/unsat.tex
dissertation v3:milestone

\chapter{Finding Unsatisfiable Cores of a Set of Polynomials using the Gr\"obner Basis Algorithm}
\label{ch:UNSAT}
Gr\"obner basis can find further applications to a vast field
of other techniques in formal verification. Satisfiability (SAT) problem is the basis of 
most decidable decision problems, as well as the basis of many formal verification techniques. 
In this chapter, we discuss a special topic branching out from SAT theory. They are about a situation 
when SAT problems give negative answer, which are called unsatisfiability (UNSAT) problems. Within 
a set of constrains (e.g. clauses, formulas or polynomials) which is unsatisfiable, sometimes 
it is worthy to explore the reasons for UNSAT. From the execution of the GB algorithm, 
an auxiliary structure can be obtained to help explore the reasons causing UNSAT. This chapter introduces the 
details about the motivation, mechanism and implementation.

\section{Motivation}
In this section, we introduce the motivation of our UNSAT core extraction research. We
start with defining an UNSAT problem, then review the previous work and applications 
to abstraction refinement. Moreover, we then present the need of finding a word-level 
analog to contemporary UNSAT reasoning techniques.

\subsection{Preliminaries of SAT/UNSAT Theory}
Boolean satisfiability (SAT) problem is one of most basic problems in computer science.
In the following part we define the terminology related to SAT problem \cite{anaICCAD}.

\begin{Definition}
A {\bf literal} $l$ is defined as a variable $v$, or its negation $\overline{v}$. A disjunction (OR relation)
of literals forms a {\bf clause}, i.e. $c = l_1\lor l_2 \lor \cdots \lor l_k$. 
A Boolean formula can always be written as {\bf conjunctive normal form} (CNF),
which is the conjunction (AND relation) of clauses: $F = c_1\land c_2\land\cdots\land c_k$.
\end{Definition}

Using above concepts, the SAT/UNSAT problem can also be formally defined:
\begin{Definition}
A {\bf satisfiability (SAT) problem} is a decision problem that takes a CNF formula and returns
that the formula is satisfiable (SAT), when there is an assignment of the variables making the 
formula evaluate to true. Otherwise, the formula is {\bf unsatisfiable (UNSAT)}.
\end{Definition}

Figure \ref{fig:SAT} shows a simple example of SAT problem on circuits verification:
we need to verify whether subcircuit $A$ and subcircuit $B$ has the same function, so
we build a miter circuit for their outputs $X$ and $Y$, and the equivalence checking 
problem is turned into a SAT problem as follows:

\begin{center}
\begin{align*}
&\textit{Is\ subcircuit\ A\ functionally\ equivalent\ to\ subcircuit\ B?}\\
&\Longleftrightarrow
\textit{Is\ it\ true\ that\ no\ Boolean\ vector\ assignment\ to\ primary\ inputs\ a,b,c\ such\ that\ Z=1?}
\end{align*}
\end{center}

@@

For an UNSAT problem, the cause of UNSAT may include a part of clauses.
\begin{Definition}
Assume a CNF formula $F$ is UNSAT. A subformula $M \subseteq F$ is an {\bf UNSAT core} 
if $M$ is also UNSAT. Further, if $\forall c\in M$, $M\setminus\{c\}$ is SAT, then 
$M$ is called a {\bf minimal UNSAT core} of $F$.
\end{Definition}

\subsection{Previous Work on CNF-UNSAT}
In most cases, SAT problems are modeled as conjunctive normal form (CNF) formulas, and solved by procedures based on 
Davis-Putnam and Davis-Logemann-Loveland (DPLL) algorithm. The main idea of DPLL algorithm
is recursive branching and backtrack searching. To improve the efficiency of DPLL-based
SAT solver, efforts are made on minimizing number of branches, accelerating unit propagation
and modifying backtrack algorithm. Recently state-of-art SAT solvers developed conflict-driven
clause learning (CDCL) technique to prune the search space, which is effective to reduce
result search time.

When a SAT solver fails to give a SAT assignment, it will provide a {\it UNSAT proof} or {\it refutation proof}
to prove the problem is UNSAT. By analyzing clauses involved to the UNSAT proof, we can 
generate a subformula which remains to be UNSAT. A n\"aive method is to collect all leaf 
clauses in an UNSAT proof as an UNSAT core. In practice, a minimal UNSAT core is more valuable.
There are mainly 2 kinds of method to find minimal UNSAT cores.
The one is insertion-based method, which is achieved by adding clauses to the smallest subset 
until the subset turns to be UNSAT. The other method is deletion-based method,
which is realized by deleting clauses from a larger subset until the subset turns to be SAT.
Recently heuristics such as clause-set refinement \cite{modelrotation} and model rotation \cite{belov2011accelerating}
are applied as improvements on the deletion/insertion based methods.
These methods are also expanded to satisfiability modulo theories (SMT) \cite{cimatti2007simple}.

On the other hand, researchers are seeking an alternative solution for SAT problem using a totally
different method from ``old-fashioned" DPLL algorithm. One promising option is polynomial calculus (PC),
mapping Boolean variables and connectors in CNF formulas to variables and operators in 
Galois fields. In this way clauses are transformed to monomials/polynomials, thus theorems and concepts in computer
algebra such as Hilbert's Nullstellensatz and Gr\"obner basis can be employed to assist finding
valid assignments or proof of unsatisfiability. Basic concepts about PC will be formally introduced with definitions
from computer algebra.

@@

The inspiration that use PC to solve SAT problems first comes from \cite{ceiSTOC96}.
Using PC and its variations, researchers develop many SAT solvers \cite{STABLE,BLUEVERI,PolyBoRi}. 
Besides, researchers borrow concepts from PC, combine them with traditional DPLL and clause learning techniques
to build hybrid SAT solvers \cite{condratTACAS07,Zengler2010}.

\subsection{Exploiting UNSAT cores for abstraction refinement}
Finding small UNSAT cores attracts interest for decades because of the need for all kinds of applications.
In solving MaxSAT problems, small UNSAT proof provides lower bound for the branch-and-bound searching algorithm \cite{li2009maxsat}.
It is applied to solving Boolean algebra problems, such as Boolean function decomposition \cite{lee2008bi}. 
Small explanation generation in general constraint programming problems also rely on small UNSAT cores \cite{cambazard2008reformulating}.

@@

The first ``if-else" branch in Algorithm \ref{alg:absrefine} can be
explained as: we check if the conjunction of formulas 
$$I(s_0)\land \bigwedge_{i=0}^{k-1}T(s_i,s_{i+1}) \land \neg p$$
is SAT or not, where $s_i$ denotes the set of reached states in $i$-th time-frame. If the result is
SAT, then a counterexample is found that violates property $p$. If the result is UNSAT, we cannot
assert that $p$ is satisfied for the original machine $M$ because we only unrolled $M$ for a given specific number of time-frames
without any fixpoint detection.
In this algorithm, we analyze the UNSAT core composed by a set of clauses whose conjunction is UNSAT.
If there are some latch variables ($L_{abs}$) not included in this UNSAT core, then we can assert that the evaluations of 
these variables will not affect the unsatisfiability of original formula. Therefore, we can ignore them in the abstracted model.
In practice, we turn these latches into primary inputs/outputs as shown in Figure \ref{fig:refine}(b) ($L_{abs} = \{l_1,\dots,l_m\}$).

@@

% {\it Gr\"obner basis is more suitable for UNSAT problems because of following theorem:}
{\it While the complexity of computing a GB over finite fields is exponential in the number of variables,
the GB  computation is observed to be more efficient for UNSAT problems.} 
The reason is as follows:
\begin{Theorem}
{\bf Weak Nullstellensatz:} Given ideal $J\subset \mathbb F[x_1,\dots,x_n]$, its variety over algebraic closure
of field $\mathbb F$ is empty if and only if its reduced Gr\"obner basis contains only one generator ``1".
$${\bf V}_{\overline{\mathbb F}}(J) = \emptyset \Longleftrightarrow \text{reduced GB}(J) = \{1\}$$
\end{Theorem}
It is well known that using Buchberger's algorithm and its variations to compute a GB has a very high space and
time complexity and is usually not practical. One reason is that the size of the GB may explode even if the term
ordering is carefully chosen. However if the reduced GB is $\{1\}$, which means every term in the original polynomials
will be canceled, the degree of remainders when computing GB with Buchberger's algorithm will be limited. 
Thus the number of polynomials in non-reduced GB is much smaller than usual.
Instead of applying polynomial calculus to SAT solving, it may be more efficient to try it for UNSAT
problems.

@@

\section{UNSAT Cores of Polynomial Ideals}
\label{sec:core}
In this section we provide a solution to the problem that how to use 
GB to extract the UNSAT core.

\textbf{Problem statement: }
Let $F = \{f_1, \dots f_s\}$ be a set of multivariate polynomials in
the ring $R = \F[x_1,\dots,x_d]$ that generate ideal $J = \langle
f_1,\dots,f_s\rangle \subset R$. Suppose that it is known that $V(J) =
\emptyset$, or it is determined to be so by applying the Gr\"obner
basis algorithm. Identify a subset of polynomials $F_c \subseteq F,
J_c = \langle F_c \rangle$, such that $V(J_c) = \emptyset$
too. Borrowing the terminology from the Boolean SAT domain, we
call $F_c$ the infeasible core or the unsat core of $F$.

@@

%% Let us first consider the problem of finding any $F_c \subset F$,
%% disregarding minimality. 
It is not hard to motivate that an unsat core should be identifiable
using the Gr\"obner basis algorithm: Assume that $F_c=F-\{f_j\}$. If
$GB(F) = GB(F_c) = \{1\}$, then it implies that $f_j$ is a member of
the ideal generated by $(F - \{f_j\})$, i.e. $f_j \in \langle F -
\{f_j\}\rangle$. Thus $f_j$ can be composed of the other
polynomials of $F_c$, so $f_j$ is  not a part of the unsat core, and
it can be safely discarded from $F_c$. This can be identified by means
of the GB algorithm for this ideal membership test.% (Definition \ref{def:gb}(ii)).

A n\"aive way (and inefficient way) to identify {\it a minimal core}
using the GB computation is as follows:
select a polynomial $f_i$ and see if $V(F_c - \{f_i\}) = \emptyset$
(i.e. if reduced $GB(F_c - \{f_i\}) = \{1\}$). If so, discard $f_i$
from the core; otherwise retain $f_i$ in $F_c$. Select a different
$f_i$ and continue until all polynomials $f_i$ are visited for
inclusion in $F_c$. This approach will produce a minimal core, as we
would have tested each polynomial $f_i$ for inclusion in the
core. This requires $O(|F|)$ calls to the GB engine, which is really
impractical.

@@

Buchberger's algorithm picks pairs of polynomials from a given set, computes their S-poly, then reduces this S-poly
with the given set of polynomials. If the remainder is non-zero,
it is added to the set of polynomials.
By tracking S-poly computations and multivariate divisions that lead to remainder 
1, we can obtain an UNSAT core. Moreover, we can identify a minimal UNSAT core with one-time
 execution of Buchberger's algorithm.

@@

\vspace{0.1in}
We compute its GB using Buchberger's algorithm with lexicographic term ordering $a>b>x>y>z$.
Since this problem is UNSAT, we will stop when ``1" is added to GB.
\begin{enumerate}[{1)}]
\item First we compute $Spoly(f_1,f_2)\xrightarrow{F}_{+} r_1$, remainder $r_1$ equals to $a$;
\item Update $F=F\cup r_1$;
\item Next we compute $Spoly(f_1,f_3)\xrightarrow{F}_{+} r_2$, remainder $r_2$ equals to $b$;
\item Update $F=F\cup r_2$;
\item We can use a directed acyclic graph (DAG) to represent the process to get $r_1,r_2$, as Figure \ref{fig:UNSAT}(a) shows;
\item Then we compute $Spoly(f_1,f_4) = s_3= a+b+1$, obviously $a+b+1$ can be reduced (multivariate divided) by
$r_1$ , the intermediate remainder $r_3 = b+1$. It can be immediately divided by $r_2$, and the remainder is ``1", we
terminate the Buchberger's algorithm;
\item We draw a DAG depicting the process through which we obtain remainder ``1" as shown in Figure \ref{fig:UNSAT}(b). 
From leaf ``1" we backtrace the graph to roots $f_1,f_2,f_3,f_4$. They constitute an UNSAT core for this problem
as these polynomials are the ``causes" of unsatisfiability of original set of polynomials.
\end{enumerate}
\end{Example}
\begin{figure}[hbt]
\centering{
\includegraphics[width=\textwidth]{newfig/UNSAT.eps}
\caption{DAG representing Spoly computations and multivariate divisions}
\label{fig:UNSAT}}
\end{figure}
% We conclude our approach as follows: i) Keep track of all $Spoly(f_i,
% f_j)\xrightarrow{f_1, \dots, f_s}_+ r$; ii) mark $f_i, f_j$ if $r\neq
% 0$; iii) Mark those $f_l$'s that have leading terms that cancel
% monomial terms in $Spoly(f_i, f_j)$; iv) Build a DAG recording all reductions including Spoly computations
% and multivariate divisions; v) Once $1\in J$ is
% detected, traverse the graph to identify all the marked polynomials that were
% used in obtaining this unit element. These polynomials constitute the
% core.

We conclude our approach as a conjecture Algorithm \ref{alg:UNSAT}.

\begin{algorithm}[hbt]
\SetAlgoNoLine
	\KwIn{A set of polynomials $F = \{f_1,f_2,\dots,f_s\}$}
	\KwOut{An UNSAT core $\{f_{m_1},f_{m_2},\dots,f_{m_t}\}$}
\Repeat{$r_l == 1$}
{
	Pick a pair $f_i,f_j\in F$ that has never been computed S-poly\;
	\If{$Spoly(f_i,f_j)\xrightarrow{F}_+ r_l \neq 0$}
	{
		$F = F\cup r_l$\;
		Create a DAG $G_l$ with $f_i,f_j$ as roots, $r_l$ as leaf, recording the S-poly, all intermediate remainders and $f_k\in F$ that cancel monomial terms in the S-poly\;
	}
}
Backward traverse the DAG for remainder ``1", replace $r_l$ with corresponding DAG $G_l$\;
\Return{All roots}
\caption {Extract UNSAT core using a variation of Buchberger's algorithm}\label{alg:UNSAT}
\end{algorithm}

@@

We investigate if it is possible to identify a core by analyzing the
$Spoly(f_i,f_j)\xrightarrow{F}_+ g_{ij}$ reductions in Buchberger's
algorithm. Since $F$ is itself an unsat core, 
definitely {\it there exists  
a sequence of Spoly reductions in Buchberger's algorithm where
$Spoly(f_i, f_j) \xrightarrow{F}_+ 1$ is achieved.} Moreover, 
polynomial reduction algorithms can be suitably modified to record
which polynomials from $F$ are used in the division leading to
$Spoly(f_i,f_j)\xrightarrow{F}_+1$. This suggests that 
we should be able to identify a core by recording
the {\it data} generated by Buchberger's algorithm --- namely, the
critical pairs($f_i,f_j$) used in the Spoly computations,
and the polynomials from $F$ used to cancel terms in the reduction
$Spoly(f_i,f_j)\xrightarrow{F}_+1$. The following example motivates
our approach to identify  $F_c \subseteq F$ using this data:

@@

Buchberger's algorithm terminates to a unique reduced GB, irrespective
of the order in which the critical pairs $(f_i,f_j)$ are selected and reduced by operation
$Spoly(f_i,f_j)\xrightarrow{F}_+g_{ij}$. Let us suppose that in the GB
computation corresponding to Example \ref{ex:1}, the first 3 critical
{\it Spoly} pairs analyzed are $(f_1, f_2), (f_3, f_4)$ and
$(f_2,f_5)$. It turns out that the Spoly-reductions corresponding to
these 3 pairs lead to the unit ideal. Recording the data
corresponding to this sequence of reductions is depicted by means of a
graph in Figure \ref{fig:refute}. We call this graph a {\it refutation tree}.

\begin{figure}[hbt]
\centering
\includegraphics[width=4.5in]{newfig/refutation_tree.eps}
\caption{Generating refutation trees to record unsat cores}
\label{fig:refute}
\end{figure}

@@

To identify an $F_c \subset F$, we start from the refutation node
``1'', and traverse the graph in reverse, all the way up to the
leaves. Then, all the leaves in the transitive fanin of ``1''
constitute an unsat core. The polynomials (nodes) that do not lie in
the transitive fanin of ``1'' can be safely discarded from $F_c$. From
Figure \ref{fig:refute}, $F_c = \{f_1,f_2,f_3,f_4,f_5\}$ is identified
as an unsat core of $F$.

\section{Reducing the Size of the Infeasible Core $F_c$}
\label{sec:alg}
The core $F_c$ obtained from the aforementioned procedure may
contain redundant elements which could be discarded. For
example, consider the core $F_c=\{f_1,\dots, f_5\}$ generated in the
previous section. While $F_c$ is a smaller infeasible core of $F$, it
is not minimal. In fact, Example 1 shows that $F_{c4} =\{f_1,f_2,f_4,f_5\}$ is the
minimal core, where $F_{c4} \subset F_{c}$. Clearly, the polynomial
$f_3 \in F_c$ is a redundant element of the core and can be 
discarded. We will now describe techniques to further reduce the size
of the unsat core by identifying such redundant elements. 
%how the size of this core could be reduced further.  
For this purpose, we perform a more 
systematic book-keeping of the data generated during the execution of
Buchberger's algorithm and the refutation tree.

\subsection{Identifying redundant polynomials from the refutation tree}

@@

{\bf Recording the coefficients:} In Equation \ref{eqn:red} we obtain a
vector of polynomial coefficients $c_k$ where $k>s$. These
coefficients are associated with new elements (remainders) in the
Gr\"obner basis, that are not a part of the unsat core. 
%which cannot benefit the UNSAT core extraction. 
Since each polynomial $f_k$, ($k>s$) is generated by
$\{f_1,\dots,f_s\}$, we can re-express $f_k$ in terms of $\{f_1,\dots,
f_s\}$. Thus, each $f_k, k>s$ can be written as $f_k = d_1f_1 + \dots
+ d_sf_s$. This process adds a new row $(d_1,\dots,d_s)$ to the
coefficient matrix $M$.

@@

\section{Iterative Refinement of the Unsat Core}
\label{sec:iter}

>>>>

\chapter{Finding Unsatisfiable Cores of a Set of Polynomials using the Gr\"obner Basis Algorithm}
\label{ch:UNSAT}
In previous chapters, we have introduced the concept of word-level abstraction and 
reachability analysis of the state-space of sequential circuits for equivalence checking.
Modern property verification techniques employ state-space abstraction using interpolants and UNSAT 
cores. In this chapter, we explore algebraic geometry analogs of UNSAT cores which can be 
applied at both bit and word-level, actually over any field.

Boolean satisfiability (SAT) problem is the basis of 
most decidable decision problems, as well as the basis of many formal verification techniques. 
In this chapter, we discuss a special topic branching out from SAT theory. It is about a situation 
when SAT problems give negative answer, which are called unsatisfiability (UNSAT) problems. Within 
a set of constrains ({\it e.g.} clauses, formulas or polynomials) which is unsatisfiable, sometimes 
it is worthy to explore the reasons for UNSAT. From the execution of the GB algorithm, 
an auxiliary structure can be obtained to help explore the reasons causing UNSAT. This chapter introduces the 
details about the motivation, mechanism and implementation.

\section{Motivation}
In this section, we introduce the motivation of our UNSAT core extraction research. We
start by defining an UNSAT problem, then review the previous work and applications 
to abstraction refinement. We then present the need for finding a word-level 
analog to contemporary UNSAT reasoning techniques.

\subsection{Preliminaries of SAT/UNSAT Theory}
Boolean satisfiability (SAT) problem is a fundamental problem in computer science.
In the following part we define the terminology related to SAT problem \cite{anaICCAD}.

\begin{Definition}
A {\bf literal} $l$ is defined as a variable $v$, or its negation $\overline{v}$. A disjunction (OR relation)
of literals forms a {\bf clause}, {\it i.e}. $c = l_1\lor l_2 \lor \cdots \lor l_k$. 
A Boolean formula can always be written as {\bf conjunctive normal form} (CNF),
which is the conjunction (AND relation) of clauses: $F = c_1\land c_2\land\cdots\land c_k$.
\end{Definition}

Using above concepts, the SAT/UNSAT problem can also be formally defined:
\begin{Definition}
A {\bf satisfiability (SAT) problem} is a decision problem that takes a CNF formula and returns
that the formula is SAT, whenever there is an assignment of variables that makes the 
formula evaluate to true. Otherwise, the formula is unsatisfiable (UNSAT).
\end{Definition}

Figure \ref{fig:SAT} shows a simple example of SAT problem on circuit verification:
we need to verify whether subcircuit $A$ and subcircuit $B$ have the same function, so
we build a miter circuit for their outputs $X$ and $Y$, and the equivalence checking 
problem is turned into a SAT problem as follows:

\vspace{-0.5in}
\begin{center}
\begin{align*}
&\textit{Is\ subcircuit\ A\ functionally\ equivalent\ to\ subcircuit\ B?}\\
&\Longleftrightarrow
\textit{Is\ it\ true\ that\ no\ Boolean\ vector\ assignment\ to\ PIs\ a,b,c\ exists\ such\ that\ Z=1?}
\end{align*}
\end{center}

@@

For an UNSAT problem, the cause of UNSAT may include a subset of clauses.
\begin{Definition}
Assume a CNF formula $F$ is UNSAT. A subformula $M \subseteq F$ is an {\bf UNSAT core} 
if $M$ is also UNSAT. Further, if $\forall c\in M$, $M\setminus\{c\}$ is SAT, then 
$M$ is called a {\bf minimal UNSAT core} of $F$.
\end{Definition}

\subsection{Previous Work on CNF-UNSAT}
In most cases, SAT problems are modeled as conjunctive normal form (CNF) formulas, and solved by procedures based on 
Davis-Putnam and Davis-Logemann-Loveland (DPLL) algorithm. The main idea of DPLL algorithm
is recursive branching and backtrack searching. To improve the efficiency of DPLL-based
SAT solver, efforts are made on minimizing number of branches, accelerating unit propagation
and modifying backtrack algorithm. Recently state-of-art SAT solvers developed conflict-driven
clause learning (CDCL) technique to prune the search space, which is effective to reduce
search time.

When a SAT solver fails to give a SAT assignment, it will provide a {\it UNSAT proof} or {\it refutation proof}
to prove the problem is UNSAT. By analyzing clauses involved in the UNSAT proof, we can 
generate a subformula which remains UNSAT. A n\"aive method is to collect all leaf 
clauses in an UNSAT proof as an UNSAT core. In practice, a minimal UNSAT core is more valuable.
There are mainly two kinds of methods to find minimal UNSAT cores.
One is insertion-based method, which is achieved by adding clauses to the smallest subset 
until the subset turns to be UNSAT. The other method is deletion-based method,
which is realized by deleting clauses from a larger subset until the subset turns to be SAT.
Recently heuristics such as clause-set refinement \cite{modelrotation} and model rotation \cite{belov2011accelerating}
are applied as improvements on the deletion/insertion based methods.
These methods are also expanded to satisfiability modulo theories (SMT) \cite{cimatti2007simple}.

On the other hand, researchers are seeking an alternative solution for SAT problem using a totally
different method from ``old-fashioned" DPLL algorithm. One promising option is polynomial calculus (PC),
mapping Boolean variables and connectors in CNF formulas to variables and operators in 
Galois fields. In this way clauses are transformed to monomials/polynomials, thus theorems and concepts in computer
algebra such as Hilbert's Nullstellensatz and Gr\"obner basis can be employed to assist finding
valid assignments or proofs of unsatisfiability. Basic concepts about PC will be formally introduced with definitions
from computer algebra.

@@

The inspiration to use PC to solve SAT problems first came from \cite{ceiSTOC96}.
Using PC and its variations, researchers developed many SAT solvers \cite{STABLE,BLUEVERI,PolyBoRi}. 
Besides, researchers borrow concepts from PC, combine them with traditional DPLL and clause learning techniques
to build hybrid SAT solvers \cite{condratTACAS07,Zengler2010}.

\subsection{Exploiting UNSAT Cores for Abstraction Refinement}
The problem of finding small UNSAT cores has attracted interest for decades because of 
its applications to various of verification and synthesis problems.
In solving MaxSAT problems, a small UNSAT proof provides lower bound for the branch-and-bound searching algorithm \cite{li2009maxsat}.
It is applied to solving logic synthesis problems, such as Boolean function decomposition \cite{lee2008bi}. 
Small explanation generation in general constraint programming problems also rely on small UNSAT cores \cite{cambazard2008reformulating}.

@@

The first ``if-else" branch in Algorithm \ref{alg:absrefine} can be
explained as: we check if the conjunction of formulas 
$$I(s_0)\land \bigwedge_{i=0}^{k-1}T(s_i,s_{i+1}) \land \neg p$$
is SAT or not, where $s_i$ denotes the set of reached states in $i$-th time-frame. If the result is
SAT, then a counterexample is found that violates property $p$. If the result is UNSAT, we cannot
assert that $p$ is satisfied for the original machine $M$ because we only unrolled $M$ for a given specific number of time-frames
without any fix-point detection.
In this algorithm, we analyze the UNSAT core composed by a set of clauses whose conjunction is UNSAT.
If there are some latch variables ($L_{abs}$) not included in this UNSAT core, then we can assert that the evaluations of 
these variables will not affect the unsatisfiability of original formula. Therefore, we can ignore them in the abstracted model.
In practice, we turn these latches into primary inputs/outputs as shown in Figure \ref{fig:refine}(b) ($L_{abs} = \{l_1,\dots,l_m\}$).

@@

% {\it Gr\"obner basis is more suitable for UNSAT problems because of following theorem:}
{\it While the complexity of computing a GB over finite fields is exponential in the number of variables,
the GB  computation is observed to be more efficient for UNSAT problems.} 
The reason is as follows:
\begin{Theorem}[Weak Nullstellensatz]
Given ideal $J\subseteq \mathbb F[x_1,\dots,x_n]$, its variety over algebraic closure $\overline{\F}$
of field $\mathbb F$ is empty if and only if its reduced Gr\"obner basis contains only one generator ``1".
$${\bf V}_{\overline{\mathbb F}}(J) = \emptyset \Longleftrightarrow \text{reduced GB}(J) = \{1\}$$
\end{Theorem}
It is well known that using Buchberger's algorithm and its variations to compute a GB has a very high space and
time complexity and is usually not practical. One reason is that the size of the GB may explode even if the term
ordering is carefully chosen. However if the reduced GB is $\{1\}$, which means every term in the original polynomials
will be canceled, the degree of remainders when computing GB with Buchberger's algorithm will be limited. 
Thus the number of polynomials in non-reduced GB is much smaller than usual.
Instead of applying polynomial calculus to SAT solving, it may be more efficient to try it for UNSAT
problems.

@@

\section{UNSAT Cores of Polynomial Ideals}
\label{sec:core}
In this section we provide a solution to the problem of how to use 
GB to extract the UNSAT core.

\textbf{Problem statement: }
Let $F = \{f_1, \dots f_s\}$ be a set of multivariate polynomials in
the ring $R = \F[x_1,\dots,x_d]$ that generate ideal $J = \langle
f_1,\dots,f_s\rangle \subset R$. Suppose that it is known that $V(J) =
\emptyset$, or it is determined to be so by applying the Gr\"obner
basis algorithm. Identify a subset of polynomials $F_c \subseteq F,
J_c = \langle F_c \rangle$, such that $V(J_c) = \emptyset$
too. Borrowing the terminology from the Boolean SAT domain, we
call $F_c$ the infeasible core or the UNSAT core of $F$.

@@

%% Let us first consider the problem of finding any $F_c \subset F$,
%% disregarding minimality. 
It is not hard to motivate that an UNSAT core should be identifiable
using the Gr\"obner basis algorithm: Assume that $F_c=F-\{f_j\}$. If
$GB(F) = GB(F_c) = \{1\}$, then it implies that $f_j$ is a member of
the ideal generated by $(F - \{f_j\})$, {\it i.e.} $f_j \in \langle F -
\{f_j\}\rangle$. Thus $f_j$ can be composed of the other
polynomials of $F_c$, so $f_j$ is  not a part of the UNSAT core, and
it can be safely discarded from $F_c$. This can be identified by means
of the GB algorithm for this ideal membership test.% (Definition \ref{def:gb}(ii)).

A n\"aive way (and inefficient way) to identify {\it a minimal core}
using the GB computation is as follows:
select a polynomial $f_i$ and see if $V(F_c - \{f_i\}) = \emptyset$
({\it i.e.} if reduced $GB(F_c - \{f_i\}) = \{1\}$). If so, discard $f_i$
from the core; otherwise retain $f_i$ in $F_c$. Select a different
$f_i$ and continue until all polynomials $f_i$ are visited for
inclusion in $F_c$. This approach will produce a minimal core, as we
would have tested each polynomial $f_i$ for inclusion in the
core. This requires $O(|F|)$ calls to the GB engine, which is really
impractical.

@@

Buchberger's algorithm picks pairs of polynomials from a given set, computes their $Spoly$, then reduces this $Spoly$
with the given set of polynomials. If the remainder is non-zero,
it is added to the set of polynomials.
By tracking $Spoly$ computations and multivariate divisions that lead to remainder 
1, we can obtain an UNSAT core. Moreover, we can identify a minimal UNSAT core with one-time
 execution of Buchberger's algorithm.

@@

\vspace{0.1in}
We compute its GB using Buchberger's algorithm with lexicographic term ordering $a>b>x>y>z$.
Since this problem is UNSAT, we will stop when ``1" is added to GB.
\begin{enumerate}[{1)}]
\item First we compute $Spoly(f_1,f_2)\xrightarrow{F}_{+} r_1$, remainder $r_1$ equals to $a$;
\item Update $F=F\cup r_1$;
\item Next we compute $Spoly(f_1,f_3)\xrightarrow{F}_{+} r_2$, remainder $r_2$ equals to $b$;
\item Update $F=F\cup r_2$;
\item We can use a directed acyclic graph (DAG) to represent the process to get $r_1,r_2$, as Figure \ref{fig:UNSAT}(a) shows;
\item Then we compute $Spoly(f_1,f_4) = s_3= a+b+1$, obviously $a+b+1$ can be reduced (multivariate divided) by
$r_1$ , the intermediate remainder $r_3 = b+1$. It can be immediately divided by $r_2$, and the remainder is ``1", we
terminate the Buchberger's algorithm;
\item We draw a DAG depicting the process through which we obtain remainder ``1" as shown in Figure \ref{fig:UNSAT}(b). 
From leaf ``1" we backtrace the graph to roots $f_1,f_2,f_3,f_4$. They constitute an UNSAT core for this problem
as these polynomials are the ``causes" of unsatisfiability of original set of polynomials.
\end{enumerate}
\end{Example}
\begin{figure}[hbt]
\centering{
\includegraphics[width=\textwidth]{newfig/UNSAT.eps}
\caption{DAG representing $Spoly$ computations and multivariate divisions}
\label{fig:UNSAT}}
\end{figure}

This general idea is depicted in Algorithm \ref{alg:UNSAT}. 
We now show a formal and efficient implementation.

\begin{algorithm}[hbt]
\SetAlgoNoLine
	\KwIn{A set of polynomials $F = \{f_1,f_2,\dots,f_s\}$}
	\KwOut{An UNSAT core $\{f_{m_1},f_{m_2},\dots,f_{m_t}\}$}
\Repeat{$r_l == 1$}
{
	Pick a pair $f_i,f_j\in F$ that has never been computed $Spoly$\;
	\If{$Spoly(f_i,f_j)\xrightarrow{F}_+ r_l \neq 0$}
	{
		$F = F\cup r_l$\;
		Create a DAG $G_l$ with $f_i,f_j$ as roots, $r_l$ as leaf, recording the $Spoly$, all intermediate remainders and $f_k\in F$ that cancel monomial terms in the $Spoly$\;
	}
}
Backward traverse the DAG for remainder ``1", replace $r_l$ with corresponding DAG $G_l$\;
\Return{All roots}
\caption {Extract UNSAT core using a variation of Buchberger's algorithm}\label{alg:UNSAT}
\end{algorithm}

@@

We investigate if it is possible to identify a core by analyzing the
$Spoly(f_i,f_j)\xrightarrow{F}_+ g_{ij}$ reductions in Buchberger's
algorithm. Since $F$ is itself an UNSAT core, 
definitely {\it there exists  
a sequence of $Spoly$ reductions in Buchberger's algorithm where
$Spoly(f_i, f_j) \xrightarrow{F}_+ 1$ is achieved.} Moreover, 
polynomial reduction algorithms can be suitably modified to record
which polynomials from $F$ are used in the division leading to
$Spoly(f_i,f_j)\xrightarrow{F}_+1$. This suggests that 
we should be able to identify a core by recording
the {\it data} generated by Buchberger's algorithm --- namely, the
critical pairs($f_i,f_j$) used in the $Spoly$ computations,
and the polynomials from $F$ used to cancel terms in the reduction
$Spoly(f_i,f_j)\xrightarrow{F}_+1$. The following example motivates
our approach to identify  $F_c \subseteq F$ using this data:

@@

Buchberger's algorithm terminates to a unique reduced GB, irrespective
of the order in which the critical pairs $(f_i,f_j)$ are selected and reduced by operation
$Spoly(f_i,f_j)\xrightarrow{F}_+g_{ij}$. Let us suppose that in the GB
computation corresponding to Example \ref{ex:1}, the first 3 critical
{\it Spoly} pairs analyzed are $(f_1, f_2), (f_3, f_4)$ and
$(f_2,f_5)$. It turns out that the $Spoly$-reductions corresponding to
these 3 pairs lead to the unit ideal. Recording the data
corresponding to this sequence of reductions is depicted by means of a
graph in Figure \ref{fig:refute}. We call this graph a {\it refutation tree}.

\begin{figure}[hbt]
\centering
\includegraphics[width=4.5in]{newfig/refutation_tree.eps}
\caption{Generating refutation trees to record UNSAT cores}
\label{fig:refute}
\end{figure}

@@

To identify an $F_c \subset F$, we start from the refutation node
``1'', and traverse the graph in reverse, all the way up to the
leaves. Then, all the leaves in the transitive fanin of ``1''
constitute an UNSAT core. The polynomials (nodes) that do not lie in
the transitive fanin of ``1'' can be safely discarded from $F_c$. From
Figure \ref{fig:refute}, $F_c = \{f_1,f_2,f_3,f_4,f_5\}$ is identified
as an UNSAT core of $F$.

\section{Reducing the Size of the UNSAT Core $F_c$}
\label{sec:alg}
The core $F_c$ obtained from the aforementioned procedure may
contain redundant elements which could be discarded. For
example, consider the core $F_c=\{f_1,\dots, f_5\}$ generated in the
previous section. While $F_c$ is a smaller infeasible core of $F$, it
is not minimal. In fact, Example 1 shows that $F_{c4} =\{f_1,f_2,f_4,f_5\}$ is the
minimal core, where $F_{c4} \subset F_{c}$. Clearly, the polynomial
$f_3 \in F_c$ is a redundant element of the core and can be 
discarded. We will now describe techniques to further reduce the size
of the UNSAT core by identifying such redundant elements. 
%how the size of this core could be reduced further.  
For this purpose, we perform a more 
systematic book-keeping of the data generated during the execution of
Buchberger's algorithm and the refutation tree.

\subsection{Identifying Redundant Polynomials from the Refutation Tree}

@@

{\bf Recording the coefficients:} In Equation \ref{eqn:red} we obtain a
vector of polynomial coefficients $c_k$ where $k>s$. These
coefficients are associated with new elements (remainders) in the
Gr\"obner basis, that are not a part of the UNSAT core. 
%which cannot benefit the UNSAT core extraction. 
Since each polynomial $f_k$, ($k>s$) is generated by
$\{f_1,\dots,f_s\}$, we can re-express $f_k$ in terms of $\{f_1,\dots,
f_s\}$. Thus, each $f_k, k>s$ can be written as $f_k = d_1f_1 + \dots
+ d_sf_s$. This process adds a new row $(d_1,\dots,d_s)$ to the
coefficient matrix $M$.

@@

\section{Iterative Refinement of the UNSAT Core}
\label{sec:iter}

As with most other UNSAT core extractors, our algorithm also cannot
generate a minimal core in one execution. To obtain a smaller core, we
 re-execute our algorithm with the core obtained in the current
iteration. We describe two heuristics that are applied to our
algorithm to increase the likelihood of generating a smaller core in
the next iteration.
